---
title: "education_analysis"
author: "HaoranLi"
date: "2026-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

This study examines:

1. What parenting styles are reported by urban Chinese parents?
2. What is the relationship between parenting styles and parent-child relationship quality?

---

# 2. Setup
```{r message=FALSE, warning=FALSE}
library(readxl)
library(dplyr)
library(psych)
library(GPArotation)
library(broom)
library(ggplot2)
library(corrplot)
library(lavaan)
library(semPlot)

```

# 3.Data import
```{r}
df <- read_excel("SPSS_data.xlsx")
df <- df %>% rename_with(trimws)
```

# 4. Data Preprocessing
## 4.1 Define Items
```{r}
ps_items <- paste0("PS-Q", 1:24)
pcr_items <- paste0("PCR-Q", 1:30)

ps_reverse <- paste0("PS-Q", 16:24)
pcr_reverse <- c("PCR-Q2","PCR-Q4","PCR-Q6","PCR-Q7","PCR-Q12","PCR-Q14","PCR-Q17","PCR-Q19","PCR-Q21","PCR-Q23","PCR-Q24","PCR-Q25")
```

## 4.2 Reverse Coding
```{r}
df2 <- df

df2[ps_reverse] <- lapply(df2[ps_reverse], function(x) 6 - x)
df2[pcr_reverse] <- lapply(df2[pcr_reverse], function(x) 6 - x)
```

## 4.3 Scale Scores
```{r}
df2 <- df2 %>%
  mutate(
    PS_score = rowMeans(select(., all_of(ps_items)), na.rm = TRUE),
    PCR_score = rowMeans(select(., all_of(pcr_items)), na.rm = TRUE)
  )
```

# 5. PCA Analysis
## 5.1 Parenting Style
```{r}
ps_data <- df2 %>% select(all_of(ps_items)) %>% na.omit()

ps_pca <- principal(ps_data, nfactors = 3, rotate = "varimax", scores = TRUE)
ps_pca
print(ps_pca$loadings, cutoff = 0.3)
```
## 5.2 Parent-Child Relationship
```{r}
pcr_data <- df2 %>% select(all_of(pcr_items)) %>% na.omit()

pcr_pca <- principal(pcr_data, nfactors = 3, rotate = "varimax", scores = TRUE)
pcr_pca
print(pcr_pca$loadings, cutoff = 0.3)
```

# 6. Factor Scores
```{r}
ps_scores <- as.data.frame(ps_pca$scores)
colnames(ps_scores) <- c("PS_authoritative","PS_authoritarian","PS_permissive")

pcr_scores <- as.data.frame(pcr_pca$scores)
colnames(pcr_scores) <- c("PCR_closeness","PCR_dependency","PCR_conflict")

complete_index <- df2 %>%
  select(all_of(c(ps_items, pcr_items))) %>%
  complete.cases()

analysis_df <- bind_cols(
  df2[complete_index, ],
  ps_scores,
  pcr_scores
)
```

# 7. Descriptive Statistics
```{r}
describe(analysis_df %>%
  select(PS_authoritative, PS_authoritarian, PS_permissive,
         PCR_closeness, PCR_dependency, PCR_conflict))
```

# 8. Correlation Analysis
```{r}
corr_data <- analysis_df %>%
  select(PS_authoritative, PS_authoritarian, PS_permissive,
         PCR_closeness, PCR_dependency, PCR_conflict)

corr_mat <- cor(corr_data, use = "pairwise.complete.obs")
corr_mat
```
```{r}
corrplot(corr_mat, method = "color",
         addCoef.col = "black",
         tl.col = "black",
         number.cex = 0.7)
```

# 9. Regression Analysis
```{r}
model_closeness <- lm(PCR_closeness ~ PS_authoritative + PS_authoritarian + PS_permissive, data = analysis_df)
model_dependency <- lm(PCR_dependency ~ PS_authoritative + PS_authoritarian + PS_permissive, data = analysis_df)
model_conflict <- lm(PCR_conflict ~ PS_authoritative + PS_authoritarian + PS_permissive, data = analysis_df)

summary(model_closeness)
summary(model_dependency)
summary(model_conflict)
```
# 10. Standardized Regression
```{r}
analysis_df_std <- analysis_df %>%
  mutate(across(everything(), scale))

model_closeness_std <- lm(PCR_closeness ~ PS_authoritative + PS_authoritarian + PS_permissive, data = analysis_df_std)
model_dependency_std <- lm(PCR_dependency ~ PS_authoritative + PS_authoritarian + PS_permissive, data = analysis_df_std)
model_conflict_std <- lm(PCR_conflict ~ PS_authoritative + PS_authoritarian + PS_permissive, data = analysis_df_std)

summary(model_closeness_std)
summary(model_dependency_std)
summary(model_conflict_std)
```
# 11. Mediation Analysis
```{r}
med_model <- '
PCR_dependency ~ a*PS_authoritative
PCR_closeness ~ cprime*PS_authoritative + b*PCR_dependency
indirect := a*b
direct := cprime
total := cprime + (a*b)
'

fit_med <- sem(med_model, data = analysis_df, se = "bootstrap", bootstrap = 2000)
summary(fit_med, standardized = TRUE, fit.measures = TRUE)
```

# 12. Path Model
```{r}
path_model <- '
PCR_closeness ~ PS_authoritative + PS_authoritarian + PS_permissive
PCR_dependency ~ PS_authoritative + PS_authoritarian + PS_permissive
PCR_conflict ~ PS_authoritative + PS_authoritarian + PS_permissive
'

fit_path <- sem(path_model, data = analysis_df)
summary(fit_path, standardized = TRUE)
```
```{r}
semPaths(
  fit_path,
  what = "std",
  layout = "tree",
  edge.label.cex = 1,
  residuals = FALSE,
  intercepts = FALSE
)
```

# 13. Conclusion
	•	Parenting styles are significantly associated with parent-child relationship quality.
	•	Authoritative parenting is positively related to closeness.
	•	Authoritarian parenting is associated with conflict.
	•	Permissive parenting shows mixed effects.
	
# 14. Limitations
	•	Cross-sectional data
	•	No causal inference
	•	Possible unobserved confounders
	
	
# Appendix
## Random forest
```{r}
library(randomForest)

rf_model <- randomForest(
  PCR_closeness ~ PS_authoritative + PS_authoritarian + PS_permissive,
  data = analysis_df,
  importance = TRUE
)

rf_model
importance(rf_model)

varImpPlot(rf_model)
```
## XGboost
```{r}
library(xgboost)

X <- as.matrix(analysis_df[, c("PS_authoritative","PS_authoritarian","PS_permissive")])
y <- analysis_df$PCR_closeness

dtrain <- xgb.DMatrix(data = X, label = y)

xgb_model <- xgboost(
  data = dtrain,
  nrounds = 100,
  objective = "reg:squarederror",
  verbose = 0
)

importance_matrix <- xgb.importance(
  feature_names = colnames(X),
  model = xgb_model
)

print(importance_matrix)
xgb.plot.importance(importance_matrix)
```
## Lasso
```{r}
library(glmnet)

X <- as.matrix(analysis_df[, c("PS_authoritative","PS_authoritarian","PS_permissive")])
y <- analysis_df$PCR_closeness

lasso_model <- cv.glmnet(X, y, alpha = 1)

plot(lasso_model)

coef(lasso_model, s = "lambda.min")
```
## Classification
```{r}
analysis_df <- analysis_df %>%
  mutate(
    closeness_high = ifelse(PCR_closeness > median(PCR_closeness), 1, 0)
  )

rf_class <- randomForest(
  as.factor(closeness_high) ~ PS_authoritative + PS_authoritarian + PS_permissive,
  data = analysis_df
)

rf_class
```


## Explaination
```{r warning=FALSE}
library(iml)

predictor <- Predictor$new(
  rf_model,
  data = analysis_df[, c("PS_authoritative","PS_authoritarian","PS_permissive")],
  y = analysis_df$PCR_closeness
)

imp <- FeatureImp$new(predictor, loss = "mse")
plot(imp)
```

